---
title: "Solving Statistical Optimization Problems with the ADMM Package"
author: Yixuan Qiu
output:
  html_document:
    toc: true
    toc_depth: 2
    theme: flatly
    highlight: default
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(eval=FALSE)
```

## Introduction

**ADMM** is an R package that utilizes the Alternating Direction Method of Multipliers
(ADMM) algorithm to solve a broad range of statistical optimization problems.
Presently the models that **ADMM** has implemented include:

- Lasso & Elastic Net
- Dantzig Selector
- Least Absolute Deviation
- Basis Pursuit

The core part of **ADMM** is written in efficient C++ code, with the help of
the [Eigen](http://eigen.tuxfamily.org) library and its R wrapper
[RcppEigen](http://cran.r-project.org/web/packages/RcppEigen/index.html).
The computational performance of **ADMM** is comparable to the most cutting-edge
R packages such as [glmnet](http://cran.r-project.org/web/packages/glmnet/index.html),
and outperforms most existing solvers on other models.

The **ADMM** package is especially suitable for large scale problems, in which
an acceptable solution can be obtained in a few iterations for a moderate
precision. It also supports parallel computing using [OpenMP](http://openmp.org)
for even bigger data.

## Installation

Currently **ADMM** can be installed from [GitHub](https://github.com/yixuan/ADMM)
using the [devtools](http://cran.r-project.org/web/packages/devtools/index.html)
package.

```{r}
library(devtools)
install_github("yixuan/ADMM")
```

**ADMM** relies on extension R packages
[RcppEigen](http://cran.r-project.org/web/packages/RcppEigen/index.html),
[rARPACK](http://cran.r-project.org/web/packages/rARPACK/index.html) and
[ggplot2](http://cran.r-project.org/web/packages/ggplot2/index.html).

## Algorithms

The **ADMM** package is an implementation of the ADMM algorithm on a number of
popular statistical models. The ADMM algorithm solves problems of the form

$$\begin{aligned}
\text{minimize}\quad & f(x)+g(z)\\
\text{subject to}\quad & Ax + Bz = c
\end{aligned}$$

where $x$, $z$ are vectors, $A$ and $B$ are matrices of suitable dimensions,
and $f$, $g$ are convex functions.

A wide range of statistical optimization problems can be written in this form,
including the ones that this package has implemented. 

ADMM algorithm can be expressed in the following iterative update equations:

$$\begin{align}
x^{k+1} & :=\underset{x}{\arg\min}\left(f(x)+\frac{\rho}{2}\Vert Ax+Bz^{k}-c+y^{k}/\rho\Vert_{2}^{2}\right)\\
z^{k+1} & :=\underset{z}{\arg\min}\left(g(z)+\frac{\rho}{2}\Vert Ax^{k+1}+Bz-c+y^{k}/\rho\Vert_{2}^{2}\right)\\
y^{k+1} & :=y^{k}+\rho(Ax^{k+1}+Bz^{k+1}-c)
\end{align}$$

where $\rho>0$ is the step size parameter. More details about this algorithm
can be found in the reference <...>.

## Models And Usages

The **ADMM** package has implemented a number of popular models in statistics and
machine learning using the algorithm introduced above. This section summarizes
the usage of various functions in **ADMM** that are related to specific models.

Unlike most other model building functions in R and extension packages, **ADMM**
makes use of the Reference Class infrastructure in R to build and fit models,
so that the syntax is in an Object-Oriented Programming (OOP) style. The typical
way to fit an model can be expressed in the following steps:

1. Call a particular model function to create a "model object".
2. Set necessary parameters and options through member functions of this model
object.
3. Actually run the estimation procedure by calling the model fitting member
function.
4. Conduct additional tasks, such as plotting and prediction.

Detailed function reference for a specific model will be given in subsequent
sections.

### Lasso

Lasso is a popular variable selection technique in high dimensional
regression analysis, which tries to find the coefficient vector $\beta$
that minimizes

$$\frac{1}{2n}\Vert y-X\beta\Vert_2^2+\lambda\Vert\beta\Vert_1$$

Here $n$ is the sample size and $\lambda$ is the regularization
parameter that controls the sparseness of $\beta$.

A Lasso model can be fit using the functions below:

- `admm_lasso(x, y, intercept, standardize, ...)`
    - This creates a model object of class `ADMM_Lasso`. It does not conduct
    the computation, but rather stores the parameters and settings of this model.
    - `x`: Predictor data matrix
    - `y`: Response Vector
    - `intercept`: Whether to fit an intercept in the model. Default is `TRUE`.
    - `standardize`: Whether to standardize the explanatory variables before
    fitting the model. Default is `TRUE`. Fitted coefficients are always
    returned on the original scale.
- `model$penalty(lambda, nlambda, lambda_min_ratio, ...)`
    - This member function sets the sequence of $\lambda$ parameters to create
    a solution path of the Lasso model. Arguments of this function have similar
    meanings as in the [glmnet](http://cran.r-project.org/web/packages/glmnet/index.html)
    package.
    - `model`: Model object, typically returned by `admm_lasso()`
    - `lambda`, `nlambda` etc.: See `?admm_lasso` for details.
- `model$parallel(nthread)`
    - This member function sets the number of threads for parallel computing.
    - `model`: Model object
    - `nthread`: Number of threads to be used
- `model$opts(maxit, eps_abs, eps_rel, rho_ratio)`
    - This member function sets options that relate to the ADMM algorithm
    - `model`: Model object
    - `maxit`: Maximum number of iterations
    - `eps_abs`: Absolute tolerance parameter
    - `eps_rel`: Relative tolerance parameter
    - `rho_ratio`: Step size parameter in the ADMM algorithm
- `model$fit()`
    - This member function starts the computation to fit the model, and returns
    an object with fields:
    - `lambda`: The sequence of $\lambda$ to build the solution path
    - `beta`: A sparse matrix containing the estimated coefficient vectors,
    each column for one $\lambda$. Intercepts are contained in the first row.
    - `niter`: Number of ADMM iterations
    - `$plot()`: Member function to create plots

The simplest syntax to obtain the model fitting result is to call

```{r}
res = admm_lasso(x, y)$fit()
```

While most member functions of class `ADMM_Lasso` return the model object itself,
it is possible (and recommended) to call a sequence of member functions in a
"chain style":

```{r}
res = admm_lasso(x, y)$penalty(nlambda = 50)$parallel(2)$opts(maxit = 100)$fit()
```

which is equivalent to

```{r}
model = admm_lasso(x, y)
model$penalty(nlambda = 50)
model$parallel(2)
model$opts(maxit = 100)
res = model$fit()
```

Clearly, the chain style syntax is more concise and intuitive.

Once the model is fitted, we can create a solution path plot by calling the
`$plot()` member function of `res`:

```{r}
res$plot()
```

### Elastic Net

### Dantzig Selector

### Least Absolute Deviation

### Basis Pursuit

## Examples