---
title: "Solving Statistical Optimization Problems with the ADMM Package"
author: Yixuan Qiu
output:
  html_document:
    toc: true
    toc_depth: 2
    theme: flatly
    highlight: default
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(eval=FALSE)
```

## Introduction

`ADMM` is an R package that utilizes the Alternating Direction Method of Multipliers
(ADMM) algorithm to solve a broad range of statistical optimization problems.
Presently the models that `ADMM` has implemented include:

- Lasso & Elastic Net
- Dantzig Selector
- Least Absolute Deviation
- Basis Pursuit

The core part of `ADMM` is written in efficient C++ code, with the help of
the [Eigen](http://eigen.tuxfamily.org) library and its R wrapper
[RcppEigen](http://cran.r-project.org/web/packages/RcppEigen/index.html).
The computational performance of `ADMM` is comparable to the most cutting-edge
R packages such as [glmnet](http://cran.r-project.org/web/packages/glmnet/index.html),
and outperforms most existing solvers on other models.

The `ADMM` package is especially suitable for large scale problems, in which
an acceptable solution can be obtained in a few iterations for a moderate
precision. It also supports parallel computing using [OpenMP](http://openmp.org)
for even bigger data.

## Installation

Currently `ADMM` can be installed from [GitHub](https://github.com/yixuan/ADMM)
using the [devtools](http://cran.r-project.org/web/packages/devtools/index.html)
package.

```{r}
library(devtools)
install_github("yixuan/ADMM")
```

`ADMM` relies on extension R packages
[RcppEigen](http://cran.r-project.org/web/packages/RcppEigen/index.html),
[rARPACK](http://cran.r-project.org/web/packages/rARPACK/index.html) and
[ggplot2](http://cran.r-project.org/web/packages/ggplot2/index.html).

## Algorithms

The `ADMM` package is an implementation of the ADMM algorithm on a number of
popular statistical models. The ADMM algorithm solves problems of the form

$$\begin{aligned}
\text{minimize}\quad & f(x)+g(z)\\
\text{subject to}\quad & Ax + Bz = c
\end{aligned}$$

where $x$, $z$ are vectors, $A$ and $B$ are matrices of suitable dimensions,
and $f$, $g$ are convex functions.

A wide range of statistical optimization problems can be written in this form,
including the ones that this package has implemented. 

ADMM algorithm can be expressed in the following iterative update equations:

$$\begin{align}
x^{k+1} & :=\underset{x}{\arg\min}\left(f(x)+\frac{\rho}{2}\Vert Ax+Bz^{k}-c+y^{k}/\rho\Vert_{2}^{2}\right)\\
z^{k+1} & :=\underset{z}{\arg\min}\left(g(z)+\frac{\rho}{2}\Vert Ax^{k+1}+Bz-c+y^{k}/\rho\Vert_{2}^{2}\right)\\
y^{k+1} & :=y^{k}+\rho(Ax^{k+1}+Bz^{k+1}-c)
\end{align}$$

where $\rho>0$ is the step size parameter. More details about this algorithm
can be found in the reference <...>.

## Models

The `ADMM` package has implemented a number of popular models in statistics and
machine learning using the algorithm introduced above. This section summarizes
the usage of various functions in `ADMM` that are related to specific models.

### Lasso

### Elastic Net

### Dantzig Selector

### Least Absolute Deviation

### Basis Pursuit

## Performance